---
title: "Tutorial Assignment 2"
output:
  pdf_document: default
  html_notebook: default
---
Sean Leggett - BDA201 February 3, 2020

## Quesion 1
Use the summary() function to produce a numerical summary of the variables in the data set

Answer:

```{r}
carsfile <- mtcars
summary(carsfile)
```

## Question 2
Use plot() function to produce a graph mpg and wt (weight). Is there a correlation between these two variables? 

Answer:
There appears to be a relationship. The lighter the car, the more efficient fuel consumption becomes.

```{r}
carsplot <- plot(carsfile$mpg, carsfile$wt,
                 main = "Car Weight vs Miles Per Gallon",
                 xlab = "Miles Per Gallon",
                 ylab = "Weight"
                 )
carsplot
```

## Question 3
Use the table() function to find out how many cars are automatic and how many are manual.

Answer:
Where mtcars$am = transmission type: 0 = Automatic, 1 = Manual
We derive 19 automatic transmissions and 13 manual transmissions.

There was no requirement to attempt formatting of the table function so I didn't attempt.

```{r}
carstable <- table(mtcars$am)
carstable
```

## Question 4
Using the subset() function separate the datasets into 2 datasets based on type of transmission.

Answer:

```{r}
autocars <- subset(carsfile, am == 0)
autocars
```

```{r}
manualcars <- subset(carsfile, am == 1)
manualcars
```

