---
title: "Weekly Assignment 2"
output:
  pdf_document: default
  html_notebook: default
---

Sean Leggett - BDA201
February 6, 2020

## Question 1
Examine the provided csv file and perform basic data inspection.

Answer:
No gaps exist in the data. The subjects have been anonymized. Fourteen fields exist. Column titles include some spaces. Last names are duplicated for Jane/John Doe.

## Question 2
What is the datatype of each feature?

Answer:
ID - Qualitative/ordinal
Last Name - Qualitative/nominal
First Name - Qualitative/nominal
City - Qualitative/nominal
State - Qualitative/nominal
Gender - Qualitative/nominal
Student Status - Qualitative/nominal
Major - Qualitative/nominal
Country - Qualitative/nominal
Age - Quantitative/interval/continuous
SAT - Quantitative/discrete
Average score (grade) - Quantitative/discrete
Height (in) - Quantitative/continuous
Newspaper readership (times/wk) - Quantitative/continuous

## Question 3
Use summary() function to display a summary of the features.

Answer:

```{r}
scores <- read.csv("Assignment_2_data.csv")
summary(scores)
```
## Question 4 and Question 5
How many males/females? How many graduate/undergraduate? Plot both using bar plots.

Answer:
Summary shows us that there are 15 each of males and females. Also, 15 each of graduates and undergraduates.

```{r}
barsplots <-
par(mfrow=c(1,2))
par(cex.axis=0.75)
barplot(table(scores$Gender))
barplot(table(scores$Student.Status))
```

## Question 6
Is the average SAT score same for graduates and undergraduates? 

Answer:
Undergraduates have a slightly higher average(mean) SAT score.1,841.2 for graduates vs 1,856.6 for undergraduates.

```{r}
grads <- subset(scores, Student.Status == "Graduate")
undergrads <- subset(scores, Student.Status == "Undergraduate")

gradsavg <- mean(grads$SAT)
underavg <- mean(undergrads$SAT)
gradsavg
underavg
```

