---
title: "Weekly Assignment 3"
output: html_notebook
---

BDA201 - Winter 2020
Sean Leggett - February 15, 2020

## Question 1  
1. A die is rolled and a coin is tossed, find the probability that the die shows an even number and
the coin shows a tail.

Answer:  
We can use the additive rule to establish the union of probablities of even dice roll and tail coin toss.

There is a base probability of rolling an even number on a fair die of 3/6 or 1 chance in 2.
There is a base probability of 1 chance in 2 resulting in tails on a coin flip.

P($A \cup B$) = P(A) + P(B) -P(A and B)
P($A \cup B$) = 3/6 +1/2 - 2/8 = 1/4

There is a 1 in 4 chance (25%) of rolling an even number and also flipping tails on a coin toss.

## Question 2  
Using R, simulate a fair 6 sided dice roll 1000 times. Generate sample space, then find and
display mean value of the 1000 randomly generated roles?

Answer:  

```{r}
sampleSpace = data.frame('Roll' = sample(1:6, 1000, replace = TRUE))
x = mean(sampleSpace$Roll)
x
```

## Question 3  
 What is the probability of getting a sum of greater than 20 when 2 fair 12-sided dice are rolled.
(Simulate this in R, by generating sample space)

Answer:  
To achieve a sum greater than 20 when rolling 2 12-sided dice, it requires a score of 11 or 12 on at least one die. Combinations that would result in a sum >29 are 10 + 11, 11 + 11, 11 +12 and 12 +12. Only 4 eventualities out of a possible 72 will satisfy the experiment.

There is a 1 in 18 chance of generating a sum greater than 20, or a 5.6% chance. To validate/demonstrate this, please see the code below. We generate a sample of numbers between 2 and 24, 1000 times, to simulate 1000 times rolling 2 x 12-sided dice.

```{r}
a <- data.frame('Die1' = sample(1:12, 1000, replace = TRUE), 'Die2' = sample(1:12, 1000, replace = TRUE))
b <- a$Die1+a$Die2
c <- b >20
d <- table(c == TRUE)
d
```

Running this code three times yields 6.1% > 20, 7.9% and 6.4% rolls greater than 20.

##Question 4  
Using the mtcars dataset, Find the probability of choosing a car at random from the dataset that
has more than 5 cylinders and has manual transmission as well.

Answer:  
