---
title: "Tutorial Acitivity 5"
output:
  pdf_document: default
  html_notebook: default
---

Sean Leggett - BDA201 - Winter 2020  
March 2, 2020  

## Question

Is there a linear relationship between the age at which a child first begins to speak and his or her mental ability later on? To answer this question a study was conducted in which the age (in months) at which a child first spoke and the child's score on an aptitude test as a teenager were recorded:

Age 14, 26, 10, 9, 15, 20, 17, 11, 8, 20

Score 96, 71, 83, 91, 99, 87, 93, 100, 104, 94

Answer the following questions using R (where applicable)

Draw a scatterplot and determine whether there appears to be a linear relationship between these two variables.
What is the relationship, how would you describe it?
Calculate and show covariance
Calculate and show Pearson correlation coefficient r
Determine and show what percentage of the variability in the aptitude score can be explained by the variability in the age at which a child begins speaking.

## Answer
Scatterplot:

```{r}
##create df
agescore <- data.frame('Age' = c(14, 26, 10, 9, 15, 20, 17, 11, 8, 20),
                       'Score' = c(96, 71, 83, 91, 99, 87, 93, 100, 104, 94)
)

agescoreplot <- plot(agescore)
agescoreplot
```

There seems to be a general correlation between early speech and later scores on mental ability. There are a couple of outliers. This is a fairly strong, negative relationship on face value; the later you learn to speak the less likely you will achieve higher scores on acuity tests. Learning to speak after the age of 25 leaves us a little suspicious of this datapoint.

Calculate Covariance  

```{r}
age <- agescore$Age
score <- agescore$Score
agecov <- cov(score, age)
agecov
```

Calculate Pearson Correlation Coefficient  

```{r}
## we prefer cor.test to simply cor
coftest <- cor.test(age, score, method = "pearson")
coftest
```
This supports the interpretation of strong negative correlation while not necessarily overwhelming.

Demonstrate what percentage of variability in aptitude scores can be explained by variability in age:  

This could be demonstrated by assuming r = 0.626839, therefore r-squared = :  

```{r}
variability <- 0.626839 * 0.626839
variability
```
39% of the variability in aptitude scores can be explained by variability in age.
