---
title: "Weekly Assignment 2"
output:
  pdf_document: default
  html_notebook: default
---

Sean Leggett - BDA201
February 6, 2020

## Question 1
Examine the provided csv file and perform basic data inspection.

Answer:
No gaps exist in the data. The subjects have been anonymized. Fourteen fields exist. Column titles include some spaces. Last names are duplicated for Jane/John Doe.

## Question 2
What is the datatype of each feature?

Answer:  
* ID - Qualitative/ordinal  
* Last Name - Qualitative/nominal  
* First Name - Qualitative/nominal  
* City - Qualitative/nominal  
* State - Qualitative/nominal  
* Gender - Qualitative/nominal  
* Student Status - Qualitative/nominal  
* Major - Qualitative/nominal  
* Country - Qualitative/nominal  
* Age - Quantitative/interval/continuous  
* SAT - Quantitative/ratio/discrete  
* Average score (grade) - Quantitative/ratio/continuous  
* Height (in) - Quantitative/ratio/continuous  
* Newspaper readership (times/wk) - Quantitative/ration/discrete  

## Question 3
Use summary() function to display a summary of the features.

Answer:

```{r}
scores <- read.csv("Assignment_2_data.csv")
summary(scores)
```
## Question 4 and Question 5
How many males/females? How many graduate/undergraduate? Plot both using bar plots.

Answer:  
Summary shows us that there are 15 each of males and females. Also, 15 each of graduates and undergraduates.

```{r}
par(mfrow=c(1,2)) ## draw layout for frames
par(cex.axis=0.75) ## smaller text on x axis to fit
barplot(table(scores$Gender))
barplot(table(scores$Student.Status))
mtext("Gender and Status Population", outer=TRUE,  cex=1, line = -.9)

```

## Question 6
Is the average SAT score same for graduates and undergraduates? 

Answer:  
Undergraduates have a slightly higher average(mean) SAT score.1,841.2 for graduates vs 1,856.6 for undergraduates.

```{r}
## subset dataframe based on status
grads <- subset(scores, Student.Status == "Graduate")
undergrads <- subset(scores, Student.Status == "Undergraduate")

## calculate means
gradsavg <- mean(grads$SAT)
underavg <- mean(undergrads$SAT)

## display
gradsavg
underavg
```

## Question 7
Between men or women? Who reads the newspaper more frequently and which group has more
variation? (show using box plots)

Answer:  
The women from this sample read more newspapers weekly than men; the men from the sample have a greater variability in number of times per week they read newspapers. The plot below illustrates this finding.

```{r}
readershipplot <- boxplot(Newspaper.readership..times.wk. ~ Gender, data = scores,
                      main = "Gender Based Weekly Newspaper Readership",
                      xlab = "Gender",
                      ylab = "Weekly Newspaper Reads")
```

## Question 8
For age, Height, and Newspaper readership, calculate the following using R:
a. Measures of location such as mean, median, and mode (where applicable).
b. Measures of variation such as variance, standard deviation and IQR

Answer:  
The code below provides the measures requested:

Part A: Mean and Median  
Calculated and grouped into dataframe for ease of presentation.
```{r}
## calculate and assign means and medians
agemean <- mean(scores$Age)
heightmean <- mean(scores$Height..in.)
readershipmean <- mean(scores$Newspaper.readership..times.wk.)
agemedian <- median(scores$Age)
heightmedian <- median(scores$Height..in.)
readershipmedian <- median(scores$Newspaper.readership..times.wk.)

## assign to dataframe for display purposes
dfLocations <- data.frame("Variable" = c("Age", "Height", "Readership"), "Mean" = c(agemean, heightmean, readershipmean), "Median" = c(agemedian, heightmedian, readershipmedian))

## display
dfLocations
```

Part A: Mode  

There is no function for mode (statistical mode) in base R. With a small dataset such as this, it is reasonable to simply sort and count.

For Age we have a dual mode of 18 and 19 each with 5 observations.
For Height we have a mode of 68 inches with 4 observations.
For Readership we have a mode of 5 with 9 observations.

There are packages to expand statistical functions and some include mode calculations. For larger datasets these would be hepful. Or else a programmatical approach leveraging some sort of count and max functions.

```{r}
agemodetest <- sort(scores$Age)
agemodetest
```

```{r}
heightmodetest <- sort(scores$Height..in.)
heightmodetest
```

```{r}
readershipmodetest <- sort(scores$Newspaper.readership..times.wk.)
readershipmodetest
```

Part B: Measures of Variation  
Calculated and grouped into dataframe for ease of presentation.
```{r}
## calculate and assign variance, SD & IQR
agevariance <- var(scores$Age)
heightvariance <- var(scores$Height..in.)
readershipvariance <- var(scores$Newspaper.readership..times.wk.)
sdage <- sd(scores$Age)
sdheight <- sd(scores$Height..in.)
sdreadership <- sd(scores$Newspaper.readership..times.wk.)
ageIQR <- IQR(scores$Age)
heightIQR <- IQR(scores$Height..in.)
readershipIQR <- IQR(scores$Newspaper.readership..times.wk.)

## combine in dataframe for presentation
dfvariation <- data.frame("Variable" = c("Age", "Height", "Readership"), "Variance" = c(agevariance, heightvariance, readershipvariance), "Standard Deviation" = c(sdage, sdheight, sdreadership), "IQR" = c(ageIQR, heightIQR, readershipIQR))

dfvariation
```

Reference Material:  
1.	Course Content  
2.	Sams Teach Yourself R in 24 Hours, Andy Nicholls, Richard Pugh, Aimee Gott. Sams, 2016.
