---
title: "Weekly Assignment 5"
output: html_notebook
---

Sean Leggett - BDA201 - Winter 2020  
March 8, 2020  

1. Is the data correlated? What does the correlation mean? What is the pearson coefficient of correlation?
Calculate by implementing the formula in r and then use the r function “cor” to verify.  
2. What is the p-value for the pearson correlation coefficient? What is meant by calculated p-value?  
3. Calculate the regression line for this data, what is the slope and y-intercept?  
4. Show the data plotted along with the regression line in brown color.  
5. What is residual for the observation (4, 7), which observation has the largest residual?  
6. What percentage of variation is explained by the regression line? What percent of variation is due to
random and unexplained factors?  

```{r}
## generate dataframe
muggings <- data.frame("Officers" = c(10, 15, 16, 1, 4, 6, 18, 12, 14, 7),
                       "Muggings" = c(5, 2, 1, 9, 7, 8, 1, 5, 3, 6)
)
muggings
```
Question 1 - Correlated?  
Let's look at some values and plots...  

```{r}
## manual calculation of Pearson coefficient of correlation
x <- muggings$Officers
y <- muggings$Muggings
X <- x - mean(x)
Y <- y - mean(y)
r <- sum(X * Y) / sqrt(sum(X*X) * sum(Y*Y))
r
## using cor() function
cor(muggings$Muggings, muggings$Officers)

```


A scatter plot:  
```{r}
mugplot <- plot(muggings$Officers, muggings$Muggings)
mugplot
```
There would appear to be a strong negative correlation. More officers = less muggings.


